---
title: "df_apply-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{df_apply-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

```{r setup, message = FALSE}
library(NeglectR)
```

## Description

The function `df_apply` is designed to apply a function to specific columns of a data frame based on a filtering function. Columns for which the filter function returns true will be modified by the function. During this process, the order of the columns is preserved.

## Usage

In the following example, we will exhibit some functionalities of `df_apply` using the `presidents` data frame.

Here is the data we will use.

```{r view-data}
presidents
```

We want to modify all logical columns so that FALSE and TRUE are replaced with the numbers zero and one, respectively. This change can be achieved by using the `as.numeric` function. In the following chunk of code, we implement this change using `df_apply`.

```{r}
df_apply(.data = presidents, .fun =  as.numeric, .filt = is.logical)
```

